The Application of Recurrent Quantification Analysis to Human-Humanoid Interaction

### 2019-09-05 docs/corrections/29-03-2019/logbook/logbook.tex
completed corrections29-03-2019

\title{Logbook for corrections }
\author{Miguel Xochicale}
\date{ \today }

@@

\begin{abstract}
This document presents a log-book for the corrections
of \emph{.../corrections/29-04-2019/comments/scanned-draft.pdf}.
\end{abstract}

@@

	\begin{verbatim}
	recommendations were added	
	\end{verbatim}
	\textit{
	SORTED:  Wed Apr  3 18:40:38 BST 2019
	}
	\\

@@

	\end{verbatim}
	\textit{
	SORTED:  
	}
	\\

@@

	\end{verbatim}
	\textit{
	SORTED:  
	}
	\\

>>>>

\title{
	Logbook of corrections \\ 
	for manuscript-draft-29-04-2019 
}
\author{Miguel Xochicale}
\date{ \today }

@@

\begin{abstract}
This is a log book document for the corrections
of \emph{.../corrections/29-04-2019/comments/scanned-draft.pdf}.
\end{abstract}

@@

	\begin{verbatim}
	recommendations were added
	(i.e. types of signals were moved to methods section)	
	\end{verbatim}
	\textit{
	SORTED:  Wed Apr  3 18:40:38 BST 2019
	}
	\\

@@

	\begin{verbatim}
	Recommendations with the use of English language were added	
	\end{verbatim}
	\textit{
	SORTED:  Tue  3 Sep 07:15:23 BST 2019
	}
	\\

	\begin{verbatim}
	You need a sentence or two here
	to explain why this is claimed.
	\end{verbatim}
	\textit{
	SORTED:  
	}
	\\

@@

	\begin{verbatim}
	Corrections with the use of English	
	\end{verbatim}
	\textit{
	SORTED:  Tue  3 Sep 07:26:41 BST 2019
	}
	\\

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Experiment}

\item  (pp. )  \_$\wedge$

	\begin{verbatim}
	Needs a statment of ethics 
	Section of Ethics has been added. However I am not sure 
	if an Appendix should be added for further informotation.
	\end{verbatim}
	\textit{
	SORTED:  Thu  5 Sep 07:14:33 BST 2019
	}
	\\

### 2019-09-05 docs/manuscript/main.tex
completed corrections29-03-2019

\section*{Conclusions}
Generally, we can conclude that using a different level of 
smoothness for time series help us to visualise and to quantify 
the variation of movements 
between participants using RSSs, RPs and RQA. 
It is important to mention that some RQA's metrics 
(e.g. DET and ENTR) are more robust to the effect of 
smoothness of time series.
However, using determined RQA metrics will be depend on 
what one want to quantify, for instance, one can compute 
predicability, organisation,  
dynamics transitions, or complexity and determinism of the 
input time series. However, RQA metrics show certain constrains with 
regard to activity type, window length and structure of the time series.
%Similarly, such differences in time series created differences in each of the 
%RQA metrics, 
For instance, RATIO and ENTR are helpful to distinguish 
differences in any of the categories of the time series (sensor, activity, 
level of smoothness and number of participant), however for certain time series
such as time series from the sensor attached to the robot seemed to have 
little variations between each of the participants. 
The latter phenomena was in a way evidently as robot degrees of freedom 
did not allow it to move with a wide range of variability.

Additionally, we can point out that even thought our experiment is 
limited to twenty healthy right-handed participants of a range age 
of mean 19.8 SD=1.39, RQA metrics show the potential to quantify 
human movement variability in the context of 
human-humanoid imitation activities.

@@

\subsection*{Inertial Sensors}
To have more fundamental understating of nature of signals collected through 
inertial sensors in the context of human-robot interaction, we are considering 
to apply derivates to the acceleration data. We can then explore the jerkiness 
of movements and therefore the nature of arm movements which typically have 
minimum jerk \cite{flash1985}, its relationship with different body 
parts \cite{devries1982, mori2012} or the application of higher derivatives 
of displacement with respect time such as snap, crackle and pop \cite{eager2016}.

\subsection*{RQA ENTR}
Having shown that RQA ENTR metrics are robust against different data
time series, we believe that further investigation is required to be done. 
For example, Marwan et al. \cite{marwan2007, marwan2015} reviewed 
different aspects to compute RPs using different criteria for neighbours, 
different norms ($L_{1-norm}$, $L_{2-norm}$, or $L_{\infty-norm}$ ) or 
different methods to select the recurrence threshold $\epsilon$, 
such as using certain percentage of the signal \cite{letellier2006}, 
the amount of noise or using a factor based on the standard deviation 
of the observational noise among many others \cite{marwan2007}.

%*******************************************************************************
%*******************************************************************************
%*******************************************************************************
\section*{Methods}
\subsection*{State Space Reconstruction}
The method of state space reconstruction was originally proposed by 
\cite{packard1980} and formalised by \cite{takens1981}. Since then, different 
investigations and disciplines have benefited from the use of the method of 
state space reconstruction \cite{aguirre2009, stergiou2011, frank2010, 
sama2013, Quintana-Duque2016}. The method of state space reconstruction is 
based on uniform time-delay embedding methodology which is a simple 
matrix implementation that can reconstruct an unknown $d-$dimensional 
manifold $M$ from a scalar time series (e.g. one-dimensional 
time series in $\mathbb{R}$). A manifold, in this context, is a 
multidimensional curved surface within a space (e.g. a saddle) 
\cite{guastello-gregson2011}.

@@

%*******************************************************************************
%*******************************************************************************
%*******************************************************************************
\section*{Experiment} \label{sec:experiment}
We conducted an experiment in the context of human-humanoid imitation (HHI) 
activities where participants were asked to imitate simple horizontal and 
vertical arm movements performed by NAO, a humanoid robot \cite{gouaillier2009}.
Such simple movements were repeated ten times for the participant 
who copied NAO's arm movements in a face-to-face imitation activity.
Also, wearable inertial measurement unit (IMU) sensors were attached 
to the right hand of the participant and to the left hand of the robot 
(Figure~\ref{fig:hri} A,C). Data were then collected with four NeMEMSi 
IMU sensors with sampling rate of 50Hz provinding tri-axial data of the 
accelerometer, gyroscope and magnetometer sensors and quaternions 
\cite{Comotti2014}. 
%A further description of the NeMEMSi IMU sensors 
%is given in \nameref{S1_AppendixB}.
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
  \centering
\includegraphics[width=1.0\textwidth]{figures/experiment/hri/pdf/hri}
    \caption{
	{\bf Human-humanoid imitation activities.} 
    		Face-to-face human-humanoid imitation (HHI) activities for 
		(A) HHI of horizontal arm movement, 
		(B) Humanoid horizontal arm movement,
		(C) HHI of vertical arm movement, and 
		(D) Humanoid vertical arm movement.
        }
    \label{fig:hri}
\end{figure}
%%---------------------------------(FIGURE)------------------------------------

>>>>

\section*{Conclusions}
Generally, we can conclude that using a different level of 
smoothness for time series help us to visualise and to quantify 
the variation of movements 
between participants using RSSs, RPs and RQA. 
It is important to mention that some RQA's metrics 
(e.g. DET and ENTR) are more robust to the effect of 
smoothness of time series.
Choice of RQA metrics will be depend on 
what to quantify, for instance, one can compute 
predicability, organisation,  
dynamics transitions, or complexity and determinism of the 
input time series. However, RQA metrics show certain constrains with 
regard to activity type, window length and structure of the time series.
For instance, RATIO and ENTR are helpful to distinguish 
differences in any of the categories of the time series (sensor, activity, 
level of smoothness and number of participant), however time series
data from the sensor attached to the robot seemed to have 
little variations between as humanoid degrees of freedom 
did not allow it to move with a wide range of variability.

Additionally, we can point out that even though our experiment is 
limited to twenty healthy right-handed participants of a range age 
of mean 19.8 SD=1.39, RQA metrics show the potential to quantify 
human movement variability in the context of 
human-humanoid imitation activities.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%	You need a sentence or two here
%	to explain why this is claimed.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\subsection*{Inertial Sensors}
To better understand the signals collected through 
inertial sensors in the context of human-humanoid interaction, 
we could apply derivates to the acceleration data. 
We can then explore the jerkiness of movements and therefore 
the nature of arm movements which typically have 
minimum jerk \cite{flash1985}, its relationship with different body 
parts \cite{devries1982, mori2012} or the application of higher derivatives 
of displacement with respect time such as snap, crackle and pop \cite{eager2016}.

\subsection*{RQA ENTR}
Having shown that RQA ENTR metrics are robust against different data
time series, we believe that further investigation is required. 
For example, Marwan et al. \cite{marwan2007, marwan2015} reviewed 
different aspects to compute RPs using different criteria for neighbours, 
different norms ($L_{1-norm}$, $L_{2-norm}$, or $L_{\infty-norm}$ ) or 
different methods to select the recurrence threshold $\epsilon$, 
such as using certain percentage of the signal \cite{letellier2006}, 
the amount of noise or using a factor based on the standard deviation 
of the observational noise among many others \cite{marwan2007}.

%*******************************************************************************
%*******************************************************************************
%*******************************************************************************
\section*{Methods}
\subsection*{State Space Reconstruction}
The method of state space reconstruction \cite{packard1980} 
\cite{takens1981} has been applied in many disciplines 
\cite{aguirre2009, stergiou2011, frank2010, 
sama2013, Quintana-Duque2016}. The method of state space reconstruction is 
based on uniform time-delay embedding methodology which is a simple 
matrix implementation that can reconstruct an unknown $d-$dimensional 
manifold $M$ from a scalar time series (e.g. one-dimensional 
time series in $\mathbb{R}$). A manifold, in this context, is a 
multidimensional curved surface within a space (e.g. a saddle) 
\cite{guastello-gregson2011}.

@@

%*******************************************************************************
%*******************************************************************************
%*******************************************************************************
\section*{Experiment} \label{sec:experiment}
We conducted an experiment in the context of human-humanoid imitation (HHI) 
activities where participants were asked to imitate simple horizontal and 
vertical arm movements performed by NAO, a humanoid robot \cite{gouaillier2009}.
Such simple movements were repeated ten times for the participant 
who copied NAO's arm movements in a face-to-face imitation activity.
Also, wearable inertial measurement unit (IMU) sensors were attached 
to the right hand of the participant and to the left hand of the robot 
(Figure~\ref{fig:hri} A,C). Data were then collected with four NeMEMSi 
IMU sensors with sampling rate of 50Hz provinding tri-axial data of the 
accelerometer, gyroscope and magnetometer sensors and quaternions 
\cite{Comotti2014}.

\subsection*{Ethics}
The experiments of this work were conducted in November 2016 and participants
confirmed reading and understanding the participant information sheet of the 
experiments and were able to withdraw from the experiment at any time without giving
any reason. The design of the experiments adhered to the University of X
regulations, data were anonymised and videos were stored only on a 
personal computer in accordance with the Data Protection Act 1998. 
%Refer to Appendix D for further information about the ethics, 
%online participation information sheets and experiment check list.

%A further description of the NeMEMSi IMU sensors 
%is given in \nameref{S1_AppendixB}.
%%---------------------------------(FIGURE)-------------------------------------
\begin{figure}[ht]
  \centering
\includegraphics[width=1.0\textwidth]{figures/experiment/hri/pdf/hri}
    \caption{
	{\bf Human-humanoid imitation activities.} 
    		Face-to-face human-humanoid imitation (HHI) activities for 
		(A) HHI of horizontal arm movement, 
		(B) Humanoid horizontal arm movement,
		(C) HHI of vertical arm movement, and 
		(D) Humanoid vertical arm movement.
        }
    \label{fig:hri}
\end{figure}
%%---------------------------------(FIGURE)------------------------------------
